---
title: "bpca, in-memory benchmarks"
author: 
  - Kasper D. Hansen
package: bpca
bibliography: bpca.bib
abstract: >
  Big PCA	
vignette: >
  %\VignetteIndexEntry{bpca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document:
    toc_float: true
---

# Packages

```{r}
suppressPackageStartupMessages(library(bpca)))
suppressPackageStartupMessages(library(irlba))
suppressPackageStartupMessages(library(matrixStats))
dataDir <- "~/Work/packages_github/bpca_data"
```

```{r describeMatrix}
describeMatrix <- function(mat) {
    cat(sprintf("class: %s\n", class(mat)[1]))
    cat(sprintf("dimension: %s x %s\n", nrow(mat), ncol(mat)))
    cat(sprintf("percent zero: %s\n", round(sum(mat == 0) / prod(dim(mat)) * 100, 1)))
    zero_rows <- sum(rowSums(mat != 0) == 0)
    cat(sprintf("number of zero rows: %s (percentage: %s)\n", zero_rows, round(zero_rows/nrow(mat) * 100, 1)))
    zero_cols <- sum(colSums(mat != 0) == 0)
    cat(sprintf("number of zero columns: %s (percentage: %s)\n", zero_cols, round(zero_cols/ncol(mat) * 100, 1)))
    invisible(NULL)
}
```
   
```{r benchmark}
benchmark <- function(...) {
    exprs <- as.list(match.call(expand.dots = FALSE)$...)
    exprsnm <- sapply(exprs, function(e) paste(deparse(e), collapse = " "))
    timings <- sapply(exprs, function(ee) {
        time <- system.time( eval(ee) )
        time
    })
    cbind(data.frame(exprs = exprsnm), t(timings)[,1:3])
}



# Datasets

```{r}
mat0 <- readRDS(file.path(dataDir, "mat0.rds"))
describeMatrix(mat0)
```

```{r}
system.time({
    out <- prcomp(t(mat0), scale = TRUE, center = TRUE)
})
```

```{r}
benchmark(
    prcomp(mat0, scale. = TRUE),
    prcomp_irlba(t(mat0), n = 10, scale. = TRUE),
    prcomp_fsvd1(t(mat0), n = 10, i = 10)
)
```

accuracy
```{r}
ref <- prcomp(t(mat0), scale. = TRUE)
ref.irlba <- prcomp_irlba(t(mat0), n = 10, scale. = TRUE)
ref.fsvd1 <- prcomp_fsvd1(mat0, n = 10, i = 10)
##
all.equal(ref$sdev[1:10], ref.irlba$sdev)
all.equal(abs(ref$rotation[,1:10]), abs(ref.irlba$rotation))
all.equal(abs(ref$x[,1:10]), abs(ref.irlba$x))
##
all.equal(ref$sdev[1:10], ref.fsvd1$sdev)
all.equal(abs(ref$rotation[,1:10]), abs(ref.fsvd1$rotation))
all.equal(abs(ref$x[,1:10]), abs(ref.fsvd1$x))
```





# irlba


```{r}
mat <- readRDS(file.path(dataDir, "mat.rds"))
mat.dge = as(t(mat), "dgeMatrix")
mat.dgC = as(t(mat), "dgCMatrix")
class(mat.dgC)
dim(mat.dge)
cmeans = colMeans(mat.dge)
csds = colSds(as.matrix(mat.dge))

system.time({
tmp <- prcomp_irlba(mat.dge, center = TRUE, scale. = TRUE, verbose = TRUE)
})

system.time({
tmp <- prcomp_irlba(mat.dgC, center = TRUE, scale. = TRUE, verbose = TRUE)
})

system.time({
tmp <- prcomp_irlba(mat.dgC, center=FALSE, scale. = FALSE, verbose = TRUE)
})

system.time({
tmp <- prcomp_irlba(mat.dgC, center=cmeans, scale. = csds, verbose = TRUE)
})

system.time({
tmp <- prcomp_irlba(mat.dge, center = TRUE, scale. = TRUE, verbose = TRUE)
})


system.time({
tmp <- prcomp_irlba(mat.dge, center = cmeans, scale. = csds, verbose = TRUE)
})

library(TENxPBMCData)
tenx = TENxPBMCData("pbmc68k")


tenx.c1 = t(as.matrix(counts(tenx[1:10000,])))
csds = colSds(tenx.c1)
cmeans = colMeans(tenx.c1)
tenx.dgC1 = as(tenx.c1, "dgCMatrix")

system.time({
    prcomp_irlba(tenx.dgC1, center = cmeans, scale. = csds, verbose = TRUE)
})






tenx.c2 = t(as.matrix(counts(tenx[20001:nrow(tenx),])))


tenx.dge = as(t(as.matrix(counts(tenx))), "dgeMatrix")
tenx.dgC = as(t(as.matrix(counts(tenx))), "dgCMatrix")


# Sessioninfo

``{r sessionInfo, results='asis', echo=FALSE}
sessionInfo()
```

# References


